\chapter{  TelApy module description}
\label{ch:TelApy_description}

As mentioned in the introduction part (section \ref{ch:intro}), the \TelApy
module is used to control the APIs of \telemacsystem in the \python programming
language. The \telemacsystem APIs are developed in \fortran. However, it is
relatively easy to use these \fortran routines directly in \python using the
"f2py" tool of the \python Scipy library \cite{Peterson2009}. This tool will
make it possible to compile \fortran code to make it accessible and usable in
\python. This compilation step is directly integrated into the compilation of
the \telemacsystem and is thus transparent to the user. Moreover, in order to
make the \TelApy module more user friendly, a \python wrapper has been developed
in order to encapsulate and simplify the different API \python calls. This set
of transformation constitutes the \TelApy module.

The first section of this chapter is dedicated to the \fortran API of
\telemacsystem. Then, the \python \TelApy module is presented.

\section{ \telemacsystem \fortran API description}
%--------------------------------------------------------------------------------
%
An API (Application Programming Interface) is a library allowing to control the
execution of a program. Here is part of the definition from Wikipedia:\\

"In computer programming, an application programming interface (API) specifies
a software component in terms of its operations, their inputs and outputs and
underlying types. Its main purpose is to define a set of functionalities that
are independent of their respective implementation, allowing both definition
and implementation to vary without compromising each other.

In addition to accessing databases or computer hardware, such as hard disk
drives or video cards, an API can be used to ease the work of programming
graphical user interface components, to allow integration of new features into
existing applications (a so-called "plug-in API"), or to share data between
otherwise distinct applications. In practice, many times an API comes in the
form of a library that includes specifications for routines, data structures,
object classes, and variables." \\

The API’s main goal is to have control on a simulation while running a case.
For example, it must allow the user to stop the simulation at any time step,
retrieve some variables values and change them. In order to make this possible,
a \fortran structure called instance is used in the API. This informatic
struture is described in the paragraph \ref{subsec:instan}. The instance
structure gives direct access to the physical memory of variables, and allows
therefore the variable control (see paragraph \ref{subsec:var}). Furthermore,
based on modifications in \telemacsystem main subroutines, it is possible to
run hydraulic case time step by time step. This will be
presented in the paragraph \ref{subsec:exec}. And finaly, the parallelism is
evoked (paragraph \ref{subsec:para}).
All \fortran routines are available in the directory "api" of \telemacsystem
sources.

\begin{WarningBlock}{Warning:}
\centering
 In the following sections, all presented API routines are related to
 \telemac{2D}. However, the API implementation of \telemacsystem modules is
 generic that is to say based on the same structure (in the following routines
 the sequence "t2d" related to \telemac{2D} module can be replace by "sis"
 related to \sisyphe module for instance).
\end{WarningBlock}

\subsection{Instanciation}
\label{subsec:instan}
An instance is an informatic structure that gathers all the variables alterable
by the API. The definition of the "instance" structure is made in a \fortran
module dedicated to this purpose and is composed of:

\begin{itemize}
\item An index defining the instance I.D.
\item A string which can contain error messages.
\item Some pointers to the concerned module variables. This is what makes it
  possible to manipulate the variables of the module by having a direct access
  to their memory location. The list of variables that can be accessed is given
  in the Table \ref{var_access}. Not all variables within \telemac{2D} are
  there. However adding a new variable is pretty easy, the 5 steps procedure is
  described in section \ref{var_add}.
\end{itemize}

\begin{table}[!H]
\begin{center}
\begin{tabular}{|cc|}
\hline
 Variable name & Definition \tabularnewline
\hline
\hline
MODEL.TIME & Current time\tabularnewline
MODEL.BCFILE & Boundary condition file \tabularnewline
MODEL.AT &  Current time\tabularnewline
MODEL.BCFILE& Boundary condition file name\tabularnewline
MODEL.BND\_TIDE& Option for tidal boundary conditions\tabularnewline
MODEL.BOTTOMELEVATION& Level of the bottom\tabularnewline
MODEL.CHESTR& Strikler on point\tabularnewline
MODEL.FAIR& Fair on point\tabularnewline
MODEL.COTE& Prescribed elevation value\tabularnewline
MODEL.CPL\_PERIOD& Coupling period with sisyphe\tabularnewline
MODEL.DEBIT& Discharge on frontier\tabularnewline
MODEL.DEBUG& Activating debug mode\tabularnewline
MODEL.FLUX\_BOUNDARIES& Flux at boundaries\tabularnewline
MODEL.GEOMETRYFILE& Name of the geomery file\tabularnewline
MODEL.METEOFILE& Name of the binary atmospheric file\tabularnewline
MODEL.FO2FILE& Name of the formatted data file 2\tabularnewline
MODEL.LIQBCFILE& Name of the liquid boundaries file\tabularnewline
MODEL.GRAPH\_PERIOD& Graphical output period\tabularnewline
MODEL.HBOR& Boundary value on h for each boundary point\tabularnewline
MODEL.IKLE& Connectivity table between element and nodes\tabularnewline
MODEL.INCWATERDEPTH& Increase in the the depth of the water\tabularnewline
MODEL.KP1BOR& Points following and preceding a boundary point\tabularnewline
MODEL.LIHBOR& Boundary type on h for each boundary point\tabularnewline
MODEL.LISTIN\_PERIOD& Listing output period\tabularnewline
MODEL.LIUBOR& Boundary type on u for each boundary point\tabularnewline
MODEL.LIVBOR& Boundary type on v for each boundary point\tabularnewline
MODEL.LT& Current time step\tabularnewline
MODEL.COMPLEO& Graphic output counter\tabularnewline
MODEL.NBOR& Global number of boundary points\tabularnewline
MODEL.NELEM& Number of element in the mesh\tabularnewline
MODEL.NELMAX& Maximum number of elements envisaged\tabularnewline
MODEL.NPOIN& Number of point in the mesh\tabularnewline
MODEL.NPTFR& Number of boundary points\tabularnewline
MODEL.NTIMESTEPS& Number of time steps\tabularnewline
MODEL.NUMLIQ& Liquid boundary numbers\tabularnewline
MODEL.POROSITY& Porosity\tabularnewline
MODEL.RESULTFILE& Name of the result file\tabularnewline
MODEL.SEALEVEL& Coefficient to calibrate sea level\tabularnewline
MODEL.TIDALRANGE& Coefficient to calibrate tidal range\tabularnewline
MODEL.UBOR& Boundary value on u for each boundary point\tabularnewline
MODEL.VBOR& Boundary value on v for each boundary point\tabularnewline
MODEL.VELOCITYU& Velocity on u\tabularnewline
MODEL.VELOCITYV& Velocity on v\tabularnewline
MODEL.WATERDEPTH& Depth of the water\tabularnewline
MODEL.X& X coordinates for each point of the mesh\tabularnewline
MODEL.XNEBOR& Normal X to 1d boundary points\tabularnewline
MODEL.Y& Y coordinates for each point of the mesh\tabularnewline
MODEL.YNEBOR& Normal Y to 1d boundary points\tabularnewline
MODEL.EQUATION& Name of the equation used\tabularnewline
\hline
\end{tabular}
\caption{Accessible variables through the API\label{var_access}}
\end{center}
\end{table}

In addition to the instance definition, the module includes all routines needed
to manipulate it (creation, deletion, and so on).

\subsection{Variable control}
\label{subsec:var}
The way in which the instance is defined (pointers) allows manipulation of
variables during the simulation. So, to get information on the variables the
following set of functions has been implemented:

\begin{itemize}

\item get the list of variables reachable with the API:
\begin{lstlisting}
subroutine get_var_list_t2d(varname, varinfo, ierr)
!
character(len=*), intent(out) :: varname(nb_var_t2d)
character(len=*), intent(out) :: varinfo(nb_var_t2d)
integer, intent(out) :: ierr
!
end subroutine
\end{lstlisting}
\begin{itemize}
\item \textbf{varname}: An array  of dimension \verb!nb_var_t2d! containing
  strings of size \verb!t2d_var_len!, which gives the list of the variable
  names.
\item \textbf{varinfo}: An array  of dimension \verb!nb_var_t2d! containing
  strings of size \verb!t2d_var_len!, which gives a short description for
  each variable.
\item \textbf{ierr}: 0 if everything went smoothly an error index otherwise.
\end{itemize}

\item get the type of a variable.

\begin{lstlisting}
subroutine get_var_type_t2d(varname, vartype,
                            readonly, ndim, ierr)
!
character(len=t2d_var_len),  intent(in)  :: varname
character(len=t2d_type_len), intent(out) :: vartype
integer,                     intent(out) :: readonly
integer,                     intent(out) :: ndim
integer,                     intent(out) :: ient
integer,                     intent(out) :: jent
integer,                     intent(out) :: kent
integer,                     intent(out) :: getpos
integer,                     intent(out) :: setpos
integer,                     intent(out) :: ierr
!
end subroutine get_var_type_t2d
\end{lstlisting}
%
\begin{itemize}
\item \textbf{varname}: The name of the variable.
\item \textbf{vartype}: The type of the variable (DOUBLE, INTEGER, STRING,
  BOOLEAN).
\item \textbf{readonly}: 0 if the variable can be only read, and 1 if it can be
  write as well.
\item \textbf{ndim}: Number of dimensions of the variable (maximum is 3 and for
  a scalar the number of dimension is 0)
\item \textbf{ient}: 1 if the index corresponds to a mesh point
\item \textbf{jent}: 1 if the index corresponds to a mesh point
\item \textbf{kent}: 1 if the index corresponds to a mesh point
\item \textbf{getpos}: give the function position after which one the variable
  information can be "get"
\item \textbf{setpos}: give the function position after which one the variable
  information can be "set"
\item \textbf{ierr}: 0 if everything went smoothly, an error index otherwise.
\end{itemize}

\item get the size of a variable.

\begin{lstlisting}
subroutine get_var_size_t2d(id, varname,
                            dim1, dim2, dim3, ierr)
!
  integer,               intent(in) :: id
  character(len=t2d_var_len), intent(in)  :: varname
  integer,               intent(out) :: dim1
  integer,               intent(out) :: dim2
  integer,               intent(out) :: dim3
  integer,               intent(out) :: ierr
!
end subroutine get_var_size_t2d
\end{lstlisting}
%
\begin{itemize}
\item \textbf{id}: The id of the instance.  
\item \textbf{varname}: The name of the variable.
\item \textbf{dim1}: Size of the first dimension.
\item \textbf{dim2}: Size of the second dimension.
\item \textbf{dim3}: Size of the third dimension.
\item \textbf{ierr}: 0 if everything went smoothly, an error index otherwise.
\end{itemize}
%
\begin{CommentBlock}{Comment:}
\centering
If the desired variable is a scalar then, it has no first, second or third
dimension and \verb!dim1!, \verb! dim2!, \verb! dim3! are equal to zero.
\end{CommentBlock}

\item get/set the value of a variable for a given index.

These functions depend on the variable type. However, the concept is the same
for all kind of types. Moreover, the type distinction is removed in TelApy based
on \python benefits. Therefore, for these reasons, the next routine header is
focused only on the double type variable. The information for the other routine
are available in the \fortran file section of API in the \telemacsystem.

\begin{lstlisting}
subroutine get_double_t2d
  (id, varname, value, global_num,
   index1, index2, index3, ierr)
!
  integer,           intent(in)  :: id
  character(len=t2d_var_len), intent(in)  :: varname
  double precision,  intent(out) :: value
  integer,           intent(in)  :: index1
  integer,           intent(in)  :: index2
  integer,           intent(in)  :: index3
  logical,           intent(in)  :: global_num
  integer,           intent(out) :: ierr
!
end subroutine get_double_t2d
\end{lstlisting}

\begin{lstlisting}
subroutine set_double_t2d
  (id, varname, value, global_num,
   index1, index2, index3, ierr)
!
  integer,           intent(in)  :: id
  character(len=t2d_var_len), intent(in)  :: varname
  double precision,  intent(in)  :: value
  integer,           intent(in)  :: index1
  integer,           intent(in)  :: index2
  integer,           intent(in)  :: index3
  logical,           intent(in)  :: global_num
  integer,           intent(out) :: ierr
!
end subroutine set_double_t2d
\end{lstlisting}

\begin{itemize}
\item \textbf{id}: the id of the instance.
\item \textbf{varname}: The name of the variable.
\item \textbf{value}: Contains the value to be read/written.
\item \textbf{index1}: Index of the first dimension (For array of at least one
  dimension, not used otherwise).
\item \textbf{index2}: Index of the second dimension (For array of at least two
  dimension, not used otherwise).
\item \textbf{index3}: Index of the third dimension (For array of at least
  three dimension, not used otherwise).
\item \textbf{global\_num}: Logical to know if the given indexes are given in
  mesh global numbering
\item \textbf{ierr}: 0 if everything went smoothly, an error index otherwise.
\end{itemize}

\end{itemize}

\subsection{Computation control}
\label{subsec:exec}
The computation control is carried out using some specific routines to launch
the simulation. These routines constitute a decomposition of the main program
of each \telemacsystem modules corresponding to the following different
computation steps:

\begin{itemize}
\item Configuration setup. This function initialises the instance and the
  listing output. The instance, characterised by the \textbf{id} integer parameter,
  represents a run of \telemac{2D}.

\begin{lstlisting}
subroutine run_set_config_t2d(id,lu,lng,ierr)
!
  integer, intent(out) :: id
  integer, intent(in)  :: lu, lng
  integer, intent(out) :: ierr
!
end subroutine run_set_config_t2d
\end{lstlisting}

\begin{itemize}
\item \textbf{id}: Contains the id of the instance.
\item \textbf{lu}: Defines the output canal for \telemac{2D} (6 will be the
  standard output).
\item \textbf{lng}: Defines the output language of \telemac{2D} (1 For French,
  2 for English).
\item \textbf{ierr}: 0 if everything went smoothly an error index otherwise.
\end{itemize}

\begin{CommentBlock}{Comment:}
\centering
In further version, the API will be able to have multiple instances running at
the same time. In the current version you can only have one instance at a time.
\end{CommentBlock}

\item Reading the \telemac{2D} steering file. This function reads the case file and
  set the variable of the \telemac{2D} steering file accordingly.

\begin{lstlisting}
subroutine run_read_case_t2d(id,cas_file, dico_file, ierr)
!
  integer, intent(out) :: id
  character(len=144), intent(in)  :: cas_file
  character(len=144), intent(in)  :: dico_file
  integer, intent(out) :: ierr
!
end subroutine run_read_case_t2d
\end{lstlisting}

\begin{itemize}
\item \textbf{id}: The id of the instance.
\item \textbf{cas\_file}: Path to the steering file.
\item \textbf{dico\_file}: Path to the \telemac{2D} dictionary.
\item \textbf{ierr}: 0 if everything went smoothly, an error index otherwise.
\end{itemize}

\begin{WarningBlock}{Warning:}
\centering
With the API we are not using the temporary folder (this folder was created by
the \python environment and all the file declared in the steering file
where copied and renamed inside that folder) which means that the name and path
given in the steering file will be used.
\end{WarningBlock}

\item Memory allocation. This function run the allocation of all the data
  needed in \telemac{2D}. Any modifications to quantities of \telemac{2D}
  should be done before the call to that function.

\begin{lstlisting}
subroutine run_allocation_t2d(id,ierr)
!
  integer, intent(in)  :: id
  integer, intent(out) :: ierr
!
end subroutine run_allocation_t2d
\end{lstlisting}

\begin{itemize}
\item \textbf{id}: The id of the instance.
\item \textbf{ierr}: 0 if everything went smoothly, an error index otherwise.
\end{itemize}

\item Initialization. This function will do the setting of the initial
  conditions of \telemac{2D} It corresponds to the time-step 0 of a
  \telemac{2D} run.

\begin{lstlisting}
subroutine run_init_t2d(id,ierr)(id,ierr)
!
  integer, intent(in)  :: id
  integer, intent(out) :: ierr
!
end subroutine run_run_init_t2d
\end{lstlisting}

\begin{itemize}
\item \textbf{id}: The id of the instance.
\item \textbf{ierr}: 0 if everything went smoothly an error index otherwise.
\end{itemize}


\item Computation function that runs one time-step of \telemac{2D}. To compute
  all time steps, a loop on this function must be done.

\begin{lstlisting}
subroutine run_timestep_t2d(id,ierr)
!
  integer, intent(in)  :: id
  integer, intent(out) :: ierr
!
end subroutine run_timestep_t2d
\end{lstlisting}


\begin{itemize}
\item \textbf{id}: The id of the instance.
\item \textbf{ierr}: 0 if everything went smoothly, an error index otherwise.
\end{itemize}

\item Finalization. This function concludes the run of \telemac{2D} and will
  deallocate all the arrays and delete the instance. To start a new execution
  of \telemac{2D} the function RUN\_SET\_CONFIG must be run again.


\begin{lstlisting}
subroutine run_finalize_t2d(id,ierr)
!
  integer, intent(in)  :: id
  integer, intent(out) :: ierr
!
end subroutine run_finalize_t2d
\end{lstlisting}

\begin{itemize}
\item \textbf{id}: The id of the instance.
\item \textbf{ierr}: 0 if everything went smoothly an error index otherwise.
\end{itemize}

\end{itemize}

For each routine defined above, the identity number of the instance is used as
an input argument allowing all computation variables to be linked with the
corresponding instance pointers. These routines are then called in the same
order to insure a correct execution of the computation in the API main
program.

\subsection{Parallelisation}
\label{subsec:para}
All steps associated with parallel computation must be performed by the user
when he chooses to launch his calculation on several processors. In this case,
after initializing the MPI environment, the user must partition the input files
(geometry file, boundary conditions files, and so on) using the \fortran
function "partel". Then, when the calculation is complete, it is necessary to
merge each subdomains result files using the "gretel" routine of the
\telemacsystem. The MPI environment can then be closed.

\section{TelApy module}

It is relatively easy to use the \fortran API routines directly in \python using
the "f2py" tool of the \python Scipy library. This tool will make it possible
to compile \fortran code such as it is accessible and usable in \python. For more
details on this tool, the interested reader can refer directly to
\cite{Peterson2009}. However, using the advantage of the \python language, it
is possible to implement a wrapper in order to provide user friendly function
of the \fortran API. Thus, a \python overlay was developed in order to
encapsulate and simplify the different API \python calls. The different \python
functions written to simplify the use of API are available in the directory
"$HOMETEL/scripts/python27/TelAPy/$".
A \doxygen documentation is available and allows the user to visualize \python
classes, functions that can be used as well as its input and output variables
and so on.

In order to launch the \doxygen documentation, the user needs to copy and paste
the link "$HOME$\-$TEL/documentation/TelAPy/doxygen/html/index.html$" into his
favorite internet browser. Then, the user can navigate in the \doxygen
environment in order to find information. For example, if the package list
widget is selected, then, all \python tools are visible such as:

\begin{itemize}
\item TelApy
\begin{itemize}
\item api
\begin{itemize}
\item api\_module: The generic \python class for \telemacsystem APIs
\item generate\_study: Automatic generator of \telemacsystem API template \python
\item hermes: Input and Output library allowing to write and read different
  \telemacsystem formats
\item sis: The \sisyphe \python class for APIs
\item t2d: The \telemac{2D} \python class for APIs
\end{itemize}
\end{itemize}
\begin{itemize}
\item tools
\begin{itemize}
\item genop: Optimization tool based on genetic algorithms
\item newop: Optimization tool based onthe SciPy minimizer function
\item polygon: Function allowing to give point indices which are in a polygon
  defined by the user
\end{itemize}
\end{itemize}
\end{itemize}

Then, if some information are desired on a specific class as for example
"api\_module", by clicking on this class, all function contained in it are listed
as follow:

\begin{itemize}
\item \verb+def _init_+: Constructor for api module
\item \verb+def set_case+: Read the steering file and run allocation
\item \verb+def init_state_default+: Initialize the state of the model Telemac
  2D with the values of discharges and water levels as indicated by the steering
  file
\item \verb+def run_one_time_step+: Run one the time step
\item \verb+def run_all_time_steps+: Run all the time steps
\item \verb+def get_mesh+: Get the 2D mesh of triangular cells
\item \verb+def get_node+: Get the nearest node number for the coordinates
  (xval, yval)
\item \verb+def get_elem+: Get the triangle where the point (xval, yval) is
\item \verb+def show_mesh+: Show the 2D mesh with topography
\item \verb+def list_variables+: List the names and the meaning of available
  variables and parameters
\item \verb+def get+: Get the value of a variable of \telemacsystem modules
\item \verb+def set+: Set the value of a variable of \telemacsystem modules
\item \verb+def get_array+: Retrieves all the values from a variable into a
  numpy array
\item \verb+def set_array+: Changes all the values from a variable into a numpy
  array
\item \verb+def get_on_polygon+: Retrieves values for point within the polygon
  poly. Warning: this works only on array that are of size NPOIN
\item \verb+def set_on_polygon+: Set a varname value on all points that are
  within the polygon poly Warning this works only on array that are of size
  NPOIN
\item \verb+def get_on_range+: Retrieves the values of the variable on the
  range given as argument
\item \verb+def set_on_range+: set the values of the variable on the range
  given as argument
\item \verb+def get_error_message+: Get the error message from the \fortran
  sources
\item \verb+def finalize+: Delete the \telemacsystem instance
\item \verb+def generate_var_info+: Returns a dictionary containing specific
  information for each variable
\end{itemize}

For each of the previous functions all information concerning the inputs and
outputs are available such as for example for the "set" fonction which allows
to get the value of the variable of \telemacsystem module:

\begin{lstlisting}[language=Python]
def set(self,varname,value,i=0,j=0,k=0,global_num=True)
\end{lstlisting}

\begin{itemize}
\item \textbf{varname}: Name of the variable
\item \textbf{i}: index otherwise.index of the first dimension
\item \textbf{j}: index otherwise.index of the second dimension
\item \textbf{k}: index otherwise.index of the third dimension
\item \textbf{global\_num}: are the index on local/global numbering variable
  value
\end{itemize}
