%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Telemac3D Documentation
%  Theory guide
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
\documentclass[Telemac3D]{../../data/TelemacDoc} % Default font size and left-justified equations

%%---------------------------------------------------------------------------
%%Equations
%%---------------------------------------------------------------------------
\renewcommand{\Grad}{\vec \nabla}
\renewcommand{\Div}{\nabla \cdot}
\renewcommand{\Lap}{\nabla^2}%{\text{lap}\hspace{0.1em}}
\newcommand{\BDiv}{\vec{\nabla} \cdot}
\newcommand{\BLap}{\vec{\nabla}^2}
\newcommand{\DivD}{\nabla_{2D} \cdot}
\newcommand{\Divast}{\nabla^* \cdot}

%%---------------------------------------------------------------------------
%% Tableaux
%%---------------------------------------------------------------------------
%\usepackage{booktabs,multirow}
%\usepackage{lscape}
%\usepackage{longtable}
%\newcommand{\minitab}[2][1]{\begin{tabular}{#1}#2\end{tabular}}
%%\usepackage{array}
%%\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
%%\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
%%\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%%---------------------------------------------------------------------------
%% Nomenclature
%%---------------------------------------------------------------------------
\usepackage{ifthen}
\usepackage{nomencl}
%
\makenomenclature
\renewcommand{\nompreamble}{\markboth{%
                            \MakeUppercase\nomname}{\MakeUppercase\nomname}%
                           }

\renewcommand{\nomgroup}[1]{
\ifthenelse{\equal{#1}{C}}{\item[\textbf{In the continuous domain}]}{
\ifthenelse{\equal{#1}{D}}{\item[\textbf{In the discrete domain}]}{
{}
}% matches discrete domain
}% matches continuous domain
}

\begin{document}

\let\cleardoublepage\clearpage
%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------
\title{TELEMAC-3D}
\subtitle{Theory guide}
\version{\telmaversion}
\author{}
\date{\today}
\maketitle
\clearpage

%----------------------------------------------------------------------------------------
%	AUTHORS PAGE
%----------------------------------------------------------------------------------------

\newpage

\thispagestyle{empty}

\chapter*{Authors and contributions}
This theory guide was created by Agnès Leroy (EDF R\&D, LNHE), based on the previous available TELEMAC-3D
documentation. In particular, most of the material came from documents written by Jean-Michel
Hervouet (EDF R\&D, LNHE), be it his book, \textit{Hydrodynamics of Free-Surface Flows}, which also lists a number of contributors, or the
TELEMAC-3D release notes. The release notes also included contributions by Chi-Tuân Pham (EDF R\&D, LNHE).
This guide, as well as the abovementioned sources, rely on previous work by Jacek Jankowski and Astrid Decoene during their respective PhD theses.
It also benefits from contributions by Lamia Abbas, Pierrick Quemar, Antoine Joly and Jacques Fontaine.

\newpage

\chapter*{Work under construction}\label{workunderconstruction}
We would like to warn the reader that this document is still under construction and
could be improved in many ways. Any remarks, suggestions, corrections, etc. are welcome
and can be submitted to the TELEMAC developing team through the forum in the Documentation
category: \url{http://www.opentelemac.org/index.php/kunena/10-documentation}.

\newpage

\chapter*{Abstract}\label{Abstract}
This guide aims at providing a comprehensive overlook of the theory behind TELEMAC-3D.
It is based on Jean-Michel Hervouet's book, \textit{Hydrodynamics of Free-Surface Flows} \cite{hervouet007}, as well as
on the previous TELEMAC-3D release notes. It is organised into four chapters: the first two describe the equations
and modelling choices in a continuous framework, as well as the principles of the Finite Elements method.
The latter was nearly directly taken from \textit{Hydrodynamics of Free Surface Flows}, but
also includes some contributions from Jacek Jankowski's PhD thesis.
After this, the time discretisation and then the space-time discretisation of the equations are described in two separate chapters.
While the first two chapters are rather classical, the last two are specific to TELEMAC-3D. They benefitted from a lot
of material from the former release notes and from \textit{Hydrodynamics of Free-Surface Flows}.
They are still unfinished at this stage. In particular, they contain some information regarding the resolution of the advection
in TELEMAC-3D and the treatment of tidal flats, but these two sections directly come from the last release notes
and have not been adapted to this guide yet. Nevertheless, we hope that these chapters will help the
TELEMAC-3D users and developers better understand what is at the core of the module.\\

The main contributions in this theory guide, compared to previous works, lie in the new format
of the document, which aims at gathering all the theoretical background of TELEMAC-3D.
It includes highlighted remarks regarding the limits and possible improvements of the models used in the code.
The chapter 3 is new: it aims at giving the reader an overview of the time scheme, even though all
the demonstrations of consistency, conservation, etc. are only valid when written for the space-time
discretised equations. New notations are adopted compared to previous works, with the hope that
they will facilitate the understanding of the numerical schemes.


%----------------------------------------------------------------------------------------
%	COPYRIGHT PAGE
%----------------------------------------------------------------------------------------

\newpage

\thispagestyle{empty}

\TelemacCopyright{}


%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS
%----------------------------------------------------------------------------------------


\pagestyle{empty} % No headers

\tableofcontents% Print the table of contents itself

%\cleardoublepage % Forces the first chapter to start on an odd page so it's on the right

\pagestyle{fancy} % Print headers again

%----------------------------------------------------------------------------------------
%	The whole thing
%----------------------------------------------------------------------------------------
%%===========================================================================
%\section{Reminders and definition of variables}
%%===========================================================================
%
%Bold variables will represent vectors and tensors, and plain font will represent scalars; i.e.:
%
%\begin{align}
%\vec{A} = \left[\begin{array}{c}
%		A_x
%		\\
%		A_y
%		\\
%		A_z
%		\end{array}\right]
%\end{align}
%
%Furthermore, the following operators are defined:
%
%\begin{subequations}
%\begin{align}
%\Grad\{A\}		= & \vec{\nabla} A
%\\
%\Grad\{\vec{A}\}		= & \vec{\nabla} \vec{A}^T
%\\
%\Div\{\vec{A}\}	= & \vec{\nabla} \cdot \vec{A}
%\\
%\Lap\{B,A\} =	& \vec{\nabla} \cdot \left( B\vec{\nabla} A\right)
%\notag\\
%			=	& \Div\{B\Grad\{A\}\}
%\\
%\BLap\{B,\vec{A}\} =	& \vec{\nabla} \cdot \left( B\vec{\nabla} \vec{A}^T\right)
%\notag\\
%						=	& \Div\{B\Grad\{\vec{A}\}\}
%\end{align}
%\end{subequations}
%
%Where $\vec{\nabla}$ is the nabla operator defined as:
%
%\begin{align}
%\vec{\nabla} = \left[\begin{array}{c}
%		\dfrac{\partial}{\partial x}
%		\\[1em]
%		\dfrac{\partial}{\partial y}
%		\\[1em]
%		\dfrac{\partial}{\partial z}
%		\end{array}\right]
%\end{align}
%
%In addition $\vec{A}^T$ is the transpose of vector $\vec{A}$.

%===========================================================================
% Nomenclature
%===========================================================================

\printnomenclature

\chapter{Equations and modelling choices in a continuous framework}\label{Chapter1}
\input{latex/Chapter1.tex}
\newpage

\chapter{Principles of the Finite Elements method}\label{Chapter2}
\input{latex/Chapter2.tex}
\newpage

\chapter{Time discretisation in TELEMAC-3D}\label{Chapter3}
\input{latex/Chapter3.tex}
\newpage

\chapter{Space-time discretisation in TELEMAC-3D}\label{Chapter4}
\input{latex/Chapter4.tex}
\newpage

\chapter{Appendix A: Thompson formulation for radiative open boundary conditions -- from \textit{Hydrodynamics of Free-Surface Flows}}\label{appendixA}
\input{latex/Appendix_thomps.tex}
\newpage

\chapter{Appendix B: Calculation of the buoyancy source terms in the transformed mesh}\label{appendixB}
\input{latex/Appendix_buoy.tex}
\newpage

%\chapter{Appendix B: Notes sur la transformation sigma}\label{appendixB}
%\input{Appendix_B.tex}
%\newpage
%
%\chapter{Appendix B: Résolution de l'équation de continuité dans le maillage transformé}\label{appendixC}
%\input{Appendix_C.tex}
%\newpage

\nomenclature[D_U*]{$\vec{U}^*$}{Fluid velocity vector in the transformed domain \dotfill (m/s)}
\nomenclature[D_U*_x]{$U^*$}{Fluid velocity along axis $x$ in the transformed domain; not that $U^*=U$ \dotfill (m/s)}
\nomenclature[D_U*_y]{$V^*$}{Fluid velocity along axis $y$ in the transformed domain; not that $V^*=V$  \dotfill (m/s)}
\nomenclature[D_U*_z]{$W^*$}{Fluid velocity along axis $z$ in the transformed domain \dotfill (m/s)}

%===========================================================================
% Bibliography
%===========================================================================

%\addcontentsline{toc}{section}{References}
\bibliographystyle{plainnat}
\bibliography{latex/telemac3d_theory_guide}

%\bibliography{../../data/biblio}

\end{document}
